---
title: "Applied machine learning: Airfoil Noise Demo"
subtitle: "Explorator Data Analysis (EDA)"
author: "Dr. Joseph P. Yurko"
date: "Fall 2021 Semester"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This RMarkdown does not provide detailed explanations, as I usually give in reports. The purpose is to demonstrate the major steps of an Exploratory Data Analysis (EDA) on a realistic data set. The data are associated with a series of experiments related to airfoil noise. The data and some basic information about the variables can be found on the [UCI Machine Learning Repository site](https://archive.ics.uci.edu/ml/datasets/airfoil+self-noise). The code within this report corresponds to my initial attempt exploring the data to give you a realistic set of steps used to explore data.  

This report mostly uses the `tidyverse`. You will also need the `visdat`, `GGally`, `corrplot`, and `corrr` packages installed to run all code chunks yourself. The `tidyverse` is loaded in the code chunk below.  

```{r, load_tidyverse_pkg}
library(tidyverse)
```

## Read data

Read in the data from the source.  

```{r, read_from_source}
data_url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00291/airfoil_self_noise.dat"

### the file has no headers and has the variables separated by tabs
df <- readr::read_tsv(data_url, col_names = FALSE)
```


Provide names for each variable.  

```{r, make_df_col_naems}
names(df) <- c("frequency", "aoa", "chord", "velocity", "displacement", "decibels")

df %>% glimpse()
```

## Basic information

Visually check for missing values using the `visdat` package.  

```{r, check_viz_miss}
visdat::vis_miss(df)
```

Check the data types visually.  

```{r, check_viz_dtypes}
visdat::vis_dat(df)
```

Check the number of unique values per variable.  

```{r, check_unique_per_variable}
df %>% purrr::map_dbl(n_distinct)
```

Two of the variables have relatively few unique values! Check the counts associated with each unique value for `chord`.  

```{r, check_chord_counts}
df %>% count(chord)
```

Check the counts associated with each unique value for `velocity`.  

```{r, check_velocity_counts}
df %>% count(velocity)
```

Check the combinations of the two.  

```{r, check_chord_vel_combo}
df %>% count(chord, velocity)
```

However, I prefer to graphically display counts with bar charts. The `chord` bar chart is below.  

```{r, viz_chord_barchart}
df %>% 
  ggplot(mapping = aes(x = as.factor(chord))) +
  geom_bar() +
  theme_bw()
```

The `velocity` bar chart is below.  

```{r, viz_velocity_barchart}
df %>% 
  ggplot(mapping = aes(x = as.factor(velocity))) +
  geom_bar() +
  theme_bw()
```

Visualize the combinations with a "dodged" bar chart.  

```{r, viz_chord_vel_barcombo}
df %>% 
  ggplot(mapping = aes(x = as.factor(chord))) +
  geom_bar(mapping = aes(fill = as.factor(velocity)),
           position = "dodge") +
  scale_fill_viridis_d("Velocity") +
  theme_bw()
```

Visualize the combinations as a heatmap.  

```{r, viz_chord_vel_heatmap}
df %>% 
  count(chord, velocity) %>% 
  ggplot(mapping = aes(x = as.factor(chord), y = as.factor(velocity))) +
  geom_tile(mapping = aes(fill = n), color = 'black') +
  geom_text(mapping = aes(label = n,
                          color = n > median(n)),
            size = 7) +
  scale_fill_viridis_c(guide = 'none', option = 'magma') +
  scale_color_manual(guide = 'none',
                     values = c("TRUE" = 'black',
                                "FALSE" = 'white')) +
  theme_bw()
```

It is important to note that even though these variables have relatively few unique values, they are **not** categorical. They are continuous and should be treated that way in models. These two have relatively few unique values due to the experimental design choice. I will refer to them as "discrete-like" or "categorical-like" to distinguish them from the other inputs which have far more unique values.  

Check the number of unique combinations of all inputs.  

```{r, check_distinct_rows}
df %>% 
  select(-decibels) %>% 
  distinct() %>% 
  dim()
```

The data set has `r nrow(df)` rows. As shown in the output display above, there are 1503 unique combinations of the 5 inputs. That means we have zero **replications** of any combination of all 5 inputs. From an experimental design perspective this is unfortunate! I would have preferred replications to allow examining the natural variability of the process.  

## Input distributions

Examine the inputs with a large number of unique values.  

```{r, viz_input_hist}
input_names <- c("frequency", "aoa", "chord", "velocity", "displacement")

df %>% 
  select(all_of(input_names)) %>% 
  select(-chord, -velocity) %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid")) %>% 
  ggplot(mapping = aes(x = value)) +
  geom_histogram(bins = 35) +
  facet_wrap(~name, scales = "free") +
  theme_bw() +
  theme(axis.text.y = element_blank())
```

Looks like angle-of-attack, `aoa`, is zero inflated. Confirm if the "spike" near zero corresponds to values at identically equal to zero.  

```{r, check_aoa_num_zero}
df %>% 
  summarise(count_aoa_equal_zero = sum(aoa == 0),
            fraction_aoa_equal_zero = mean(aoa == 0))
```

The `frequency` variable looks highly right skewed.  

```{r, check_freq_bounds}
df %>% select(frequency) %>% summary()
```

Apply the log10 transformation to `frequency`. We saw previouisly that `frequency` has 21 unique values. The histogram below uses 35 bins which results in "gaps" appearing between the bins. This highlights what appears to be a grid like structure within the log10 space of `frequency`.  

```{r, viz_log10_freq_hist}
df %>% 
  ggplot(mapping = aes(x = log10(frequency))) +
  geom_histogram(bins = 35) +
  theme_bw()
```

Confirm the grid-like structure by including the rug "tick" marks and using just 21 bins.  

```{r, viz_log10_freq_hist_b}
df %>% 
  ggplot(mapping = aes(x = log10(frequency))) +
  geom_rug(color = 'red', size = 2) +
  geom_histogram(bins = 21) +
  theme_bw()
```

The `displacement` variable also looks highly right skewed. Check the bounds and summary stats on the `displacement` variable.  

```{r, check_displ_stats}
df %>% select(displacement) %>% summary()
```

Apply the natural log transformation to `displacement`. The histogram of the transformed variable with the rug marks is shown below. We know from before that `displacement` has over 100 unique values.  

```{r, viz_logdisp_hist}
df %>% 
  ggplot(mapping = aes(x = log(displacement))) +
  geom_rug(color = 'red', size = 1, alpha = 0.2) +
  geom_histogram(bins = 35) +
  theme_bw()
```

Apply the transformations above and remove the original variables for simplicity.  

```{r, change_df}
df_b <- df %>% 
  mutate(log_d = log(displacement),
         log_f = log10(frequency)) %>% 
  select(-displacement, -frequency)
```


Look at the distribution of the more continuous like inputs for each combination of the "discrete-like" inputs. We saw previously in the heatmap that some combinations of `chord` and `velocity` have as few as 30 observations. Because of that, use a kernel density estimate with a bandwidth encouraged to smooth the distribution more than the default settings. Visualize the transformed `displacement` and `frequency` variables instead of the original variables.  

```{r, viz_input_hist_by_group}
df_b %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = c("aoa", "log_d", "log_f")) %>% 
  ggplot(mapping = aes(x = value)) +
  geom_rug(color = 'red') +
  geom_density(mapping = aes(group = interaction(name, chord, velocity),
                             color = as.factor(chord)),
               size = 1.2, adjust = 1.35) +
  facet_grid(velocity ~ name, labeller = "label_both", scales = "free") +
  scale_color_viridis_d("chord", option = 'inferno') +
  theme_bw() +
  theme(panel.background = element_rect(fill = 'grey70'),
        panel.grid.minor = element_blank())
```

Focus on the distribution of angle-of-attack, `aoa`, since it was difficult to see in the figure above.  

```{r, viz_input_aoa_by_group}
df_b %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = c("aoa", "log_d", "log_f")) %>% 
  filter(name == 'aoa') %>% 
  ggplot(mapping = aes(x = value)) +
  geom_rug(color = 'red') +
  geom_density(mapping = aes(group = interaction(name, chord, velocity),
                             color = as.factor(chord)),
               size = 1.2, adjust = 1.35) +
  facet_grid(velocity ~ name, labeller = "label_both", scales = "free") +
  scale_color_viridis_d("chord", option = 'inferno') +
  theme_bw() +
  theme(panel.background = element_rect(fill = 'grey70'),
        panel.grid.minor = element_blank())
```

Examine the summary statistics associated with the continuous-like inputs for each "discrete-like" combination. The summary statistics are visualized with boxplots below. The angle-of-attack `aoa` is decreasing as `chord` increases~ The `log_d` becomes more concentrated near its center as `chord` increases.  

```{r, viz_input_box_by_group}
df_b %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = c("aoa", "log_d", "log_f")) %>% 
  ggplot(mapping = aes(x = as.factor(chord), y = value)) +
  geom_boxplot(mapping = aes(fill = as.factor(velocity),
                             color = as.factor(velocity)),
               alpha = 0.35) +
  facet_grid(name ~ ., scales = "free_y") +
  scale_fill_viridis_d("Velocity") +
  scale_color_viridis_d("Velocity") +
  theme_bw()
```

Pairs plot between all inputs.  

```{r, viz_pairs_inputs_all}
df_b %>% 
  select(-decibels) %>% 
  GGally::ggpairs(progress = FALSE,
                  diag = list(continuous = GGally::wrap('barDiag', bins=25))) +
  theme_bw()
```

The pairs plot allows us to see which input pairs have a "grid-like" structure in this application. For example, the `chord` and `velocity` variables have a clear "grid-like" pattern. We already looked at that pattern with the heatmap previously. We can look at those same counts with a "count scatter plot", as shown below.  

```{r, viz_chord_vel_count_scatter}
df_b %>% 
  ggplot(mapping = aes(x = chord, y = velocity)) +
  geom_count() +
  scale_size_continuous(range = c(1, 10)) +
  theme_bw()
```

The `chord` and `aoa` also have a clear structure where the `aoa` range depends on the value of `chord`.  

```{r, viz_chord_aoa_scatter}
df_b %>% 
  ggplot(mapping = aes(x = aoa, y = chord)) +
  geom_point(size = 4.5) +
  theme_bw()
```

The count plot below helps us see which combinations have high vs low counts.  

```{r, viz_chord_aoa_scatter_count}
df_b %>% 
  ggplot(mapping = aes(x = aoa, y = chord)) +
  geom_count() +
  scale_size_continuous(range = c(1, 10)) +
  theme_bw()
```

Break up by `velocity`.  

```{r, viz_chord_aoa_scatter_count_b}
df_b %>% 
  ggplot(mapping = aes(x = aoa, y = chord)) +
  geom_count() +
  facet_wrap(~velocity, labeller = "label_both") +
  theme_bw()
```

Next, let's focus on the scatter plot between `log_f` and `log_d`.  

```{r, viz_logf_logd_scatter}
df_b %>% 
  ggplot(mapping = aes(x = log_f, y = log_d)) +
  geom_point() +
  theme_bw()
```

Check the counts for each combination.  

```{r, viz_logf_logd_scatter_count}
df_b %>% 
  ggplot(mapping = aes(x = log_f, y = log_d)) +
  geom_count() +
  theme_bw()
```

Looks like there are either 1 or 2 observations per combination. Check below.  

```{r, viz_logf_logd_scatter_counb}
df_b %>% 
  count(log_f, log_d) %>% 
  ggplot(mapping = aes(x = log_f, y = log_d)) +
  geom_point(mapping = aes(color = as.factor(n),
                           size = n)) +
  scale_color_brewer("n", palette = "Set1") +
  scale_size_continuous(guide = 'none') +
  theme_bw()
```

Breakup by the `chord` and `velocity` combinations and color by `aoa`. The graphic below is visualizing all 5 inputs variables together in a single chart!  

```{r, viz_all_inputs_together}
df_b %>% 
  ggplot(mapping = aes(x = log_f, y = log_d)) +
  geom_point(mapping = aes(color = aoa)) +
  facet_grid(velocity ~ chord, labeller = "label_both") +
  scale_color_viridis_c() +
  theme_bw()
```

The pairs plot had revealed the rather high correlation between `aoa` and `log_d`. Let's check below.  

```{r, viz_pairs_inputs_continuous}
df_b %>% 
  select(aoa, log_d, log_f) %>% 
  GGally::ggpairs(progress = FALSE,
                  diag = list(continuous = GGally::wrap('barDiag', bins=25))) +
  theme_bw()
```

If we just want to focus on the correlation coefficient between variables we can use a correlation plot.  

```{r, viz_input_corrplot}
df_b %>% 
  select(-decibels) %>% 
  cor() %>% 
  corrplot::corrplot(method = 'number', type = 'upper')
```

Let's focus on the relationship between `aoa` and `log_d`.  

```{r, viz_aoa_logd_scatter}
df_b %>% 
  ggplot(mapping = aes(x = aoa, y = log_d)) +
  geom_point(size = 4.5) +
  theme_bw()
```

Break up by the `chord` and `velocity` combinations. The relationship between the two is now very clear! The log transformed `displacement` appears nearly linearly related to the angle-of-attack.

```{r, viz_aoa_logd_combos_scatter}
df_b %>% 
  ggplot(mapping = aes(x = aoa, y = log_d)) +
  geom_point(size = 2) +
  facet_grid(velocity ~ chord, labeller = "label_both") +
  theme_bw()
```

Let's check the counts.  

```{r, viz_aoa_logd_combos_scatter_count}
df_b %>% 
  ggplot(mapping = aes(x = aoa, y = log_d)) +
  geom_count() +
  facet_grid(velocity ~ chord, labeller = "label_both") +
  theme_bw()
```

Let's check the relationship between the original `displacement` variable and the angle-of-attack. As shown below, we still have a relationship between the variables. That relationship just appears exponential like in the original `displacement` space.  

```{r, viz_aoa_displ_scatter}
df %>% 
  ggplot(mapping = aes(x = aoa, y = displacement)) +
  geom_point(size = 2) +
  facet_grid(velocity ~ chord, labeller = "label_both") +
  theme_bw()
```

We saw previously that there are no replications of the 5 inputs in the data set. Since the `aoa` and log-transformed `displacement` appear so highly related to each other let's check if there are replications when we remove `log_d` from the variable set. As shown below there are no replications when we consider just four out of the five inputs!  

```{r, check_reps_wl_log_d}
df_b %>% 
  select(chord, velocity, log_f, aoa) %>% 
  distinct() %>% 
  nrow()
```

This is representative of the high correlation between `aoa` and `log_d`. Let's do one final check by calculating the correlation coefficient between the "continuous-like" inputs for each `velocity` and `chord` combination. The correlation coefficient between `aoa` and `log_d` for each `chord` and `velocity` group can be essentially 1!  

```{r, viz_corr_by_grouping}
df_b %>% 
  select(-decibels) %>% 
  group_by(chord, velocity) %>% 
  tidyr::nest() %>% 
  mutate(cor_wf = map(data, corrr::correlate, quiet = TRUE, diagonal = 1),
         cor_lf = map(cor_wf, corrr::stretch)) %>% 
  select(chord, velocity, cor_lf) %>% 
  tidyr::unnest(cor_lf) %>% 
  ungroup() %>% 
  ggplot(mapping = aes(x = x, y = y)) +
  geom_tile(mapping = aes(fill = r),
            color = 'white') +
  geom_text(mapping = aes(label = round(r, 2),
                          color = abs(r) > 0.75),
            size = 2.5) +
  facet_grid(velocity ~ chord, labeller = "label_both") +
  scale_fill_gradient2('corr',
                       low = 'red', mid = 'white', high = 'blue',
                       midpoint = 0,
                       limits = c(-1, 1)) +
  scale_color_manual(guide = 'none',
                     values = c("TRUE" = 'white',
                                "FALSE" = 'black')) +
  labs(x='', y='') +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90, vjust=0.5, size=4.5),
        axis.text.y = element_text(size = 5.5))
```

## Output visualizations

The previous figures focused on the inputs. Let's now consider the output (response) and the relationships between the output and the inputs.  

Histogram of the response with quartiles and rug marks included for reference.  

```{r, viz_output_hist}
df_b %>% 
  ggplot(mapping = aes(x = decibels)) +
  geom_rug(alpha = 0.2) +
  geom_histogram(bins = 41) +
  geom_vline(xintercept = quantile(df_b$decibels),
             color = 'red', size = 1.2, linetype = 'dashed') +
  theme_bw()
```

Visualize the output distribution for each `chord` and `velocity` combination. Use a kernel density estimate since the number of observations for some of the groups is relatively low.  

```{r, viz_output_hist_group}
df_b %>% 
  ggplot(mapping = aes(x = decibels)) +
  geom_density(mapping = aes(color = as.factor(velocity)),
               size = 1.1,
               adjust = 1.35) +
  facet_wrap(~chord, labeller = "label_both") +
  scale_color_viridis_d("Velocity") +
  theme_bw()
```

Visualize the key summary statistics per `chord` and `velocity` combination with boxplots.  

```{r, viz_output_box_groups}
df_b %>% 
  ggplot(mapping = aes(x = as.factor(chord), y = decibels)) +
  geom_boxplot(mapping = aes(group = interaction(chord, velocity),
                             fill = as.factor(velocity),
                             color = as.factor(velocity)),
               alpha = 0.35) +
  scale_fill_viridis_d("Velocity") +
  scale_color_viridis_d("Velocity") +
  theme_bw()
```

Include the mean per group and the $\pm$ 2 standard error interval on the mean.  

```{r, viz_output_box_groups_w_mean}
df_b %>% 
  ggplot(mapping = aes(x = as.factor(chord), y = decibels)) +
  geom_boxplot(mapping = aes(group = interaction(chord, velocity),
                             fill = as.factor(velocity),
                             color = as.factor(velocity)),
               alpha = 0.2) +
  stat_summary(fun.data = "mean_se",
               mapping = aes(group = interaction(chord, velocity),
                             color = as.factor(velocity)),
               fun.args = list(mult = 2),
               position = position_dodge(0.75)) +
  scale_fill_viridis_d("Velocity") +
  scale_color_viridis_d("Velocity") +
  theme_bw()
```

Look at the scatter plot between `aoa` and `log_f` where the marker color corresponds to the response value. Use separate facets for each `chord` and `velocity` combination.  

```{r, viz_output_surf_per_group}
df_b %>% 
  ggplot(mapping = aes(x = aoa, y = log_f)) +
  geom_point(size = 3.5,
             mapping = aes(color = decibels)) +
  facet_grid(velocity ~ chord, labeller = "label_both") +
  scale_color_viridis_c() +
  theme_bw()
```

Summarize the response value in case of overlap in the previous figure. Let the x-axis scales vary by facet column. Figure is tough to look at!  

```{r, viz_output_surf_per_group_mean}
df_b %>% 
  ggplot(mapping = aes(x = aoa, y = log_f)) +
  stat_summary_2d(fun = 'mean',
                  mapping = aes(z = decibels)) +
  facet_grid(velocity ~ chord, labeller = "label_both", scales = "free_x") +
  scale_fill_viridis_c() +
  theme_bw()
```

Instead, plot the response with respect to `log_f` for each `chord` and `velocity` combination colored by `aoa`. There seems to be non-linear trends between the response and `log_f`! Those trends appear to be depend on the angle-of-attack, and the `chord`!  

```{r, viz_response_logf_scatter_group}
df_b %>% 
  ggplot(mapping = aes(x = log_f, y = decibels)) +
  geom_point(mapping = aes(color = aoa)) +
  facet_grid(velocity ~ chord, labeller = "label_both") +
  scale_color_viridis_c(option = 'plasma') +
  theme_bw()
```

Include a non-linear smoother to help make it easier to see the trends.  

```{r, viz_response_logf_smooth_group}
df_b %>% 
  ggplot(mapping = aes(x = log_f, y = decibels)) +
  geom_point(mapping = aes(color = aoa), alpha = 0.5) +
  geom_smooth(formula = y ~ x,
              mapping = aes(group = interaction(chord, velocity, aoa),
                            color = aoa), 
              size = 1.) +
  facet_grid(velocity ~ chord, labeller = "label_both") +
  scale_color_viridis_c(option = 'plasma') +
  theme_bw()
```

